---
title: "Analysis of Crime Dataset using R"
author: "Thejas Raju"
date: "May 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TASK -1: Data Manipulation
### 1) Loading the Dataset EurostatCrime2015.csv 
```{r}
df <- read.csv("EurostatCrime2015.csv",
               , row.names = 1, header = TRUE)
```
### 2) Size and structure of the dataset
```{r}
size<-dim(df)#Displays Number of rows and number of columns
sprintf("Number of Rows: %i Number of Columns: %i",size[1],size[2])
structure(df)
``` 
### 3)i) Adding new column Sex.crime
```{r}
df$Sex.crime<- df$Rape+df$Sexual.assault+df$Sexual.violence
colnames(df)
```
### 3)ii) Removing columns Rape, Sexual.assault and Sexual.violence
```{r}
df[,c('Rape','Sexual.assault','Sexual.violence')]<-NULL
colnames(df)
```
### 4) Listing countries with 'na' Values
```{r}
countriesWNA <- df[rowSums(is.na(df)) > 0,]
rownames(countriesWNA)
```
### 5) Removing countries with missing data
```{r}
df_new<-na.omit(df)
df_new
```
### 6) Size of the new dataframe
```{r}
size<-dim(df_new) #Displays number of rows followed by number of columns
sprintf("Number of Rows: %i Number of Columns: %i",size[1],size[2])
```
# TASK-2: Data Analysis

### 1) Most Common Crime in Ireland in 2015
```{r}
infoIreland <- df_new['Ireland',]
commonCrime<-colnames(infoIreland)[apply(infoIreland,1,which.max)]
sprintf("Most Common Crime in Ireland is: %s", commonCrime)
```
### 2) Three most least common crimes in Ireland in 2015
```{r}
sorted <- sort(infoIreland)
leastCrimes <- data.frame(c(sorted[1:3]))
lcrimes<-colnames(leastCrimes)
sprintf("The three most least crimes in Ireland are: %s, %s, and %s",lcrimes[1],lcrimes[2],lcrimes[3])
```
### 3) Country with the highest record of offences
```{r}
df_new$highestOffences <- df_new$Assault+ df_new$Intentional.homicide + df_new$Robbery + df_new$Theft + df_new$Sex.crime
max_crime<-max(df_new$highestOffences)
name<-rownames(df_new)[df_new$highestOffences == max_crime]
sprintf("The country with highest record of offences is: %s", name)
```

#TASK-3: Data Visualisation
## The following code generates a map of the European countries coloured with a gradient that represents the average crime rate of the particular country
```{r}
library(ggplot2)
library(grid)
library(sp)
library(rworldmap)
library(maps)
library(mapproj)

wMap <- getMap()

countries <- rownames(df)

findEU <- which(wMap$NAME%in%countries)

Coords <- lapply(findEU, function(i){
  df <- data.frame(wMap@polygons[[i]]@Polygons[[1]]@coords)
  df$region =as.character(wMap$NAME[i])
  colnames(df) <- list("long", "lat", "region")
  return(df)
})

Coords <- do.call("rbind", Coords)
value=rowMeans(df, na.rm=FALSE)
df2<- data.frame(country=countries, value= value)
Coords$value <-df2$value[match(Coords$region ,df2$country)]

graph <- ggplot() + geom_polygon(data = Coords, aes(x = long, y = lat, group = region, fill = value),colour = "black", size = 0.5) +coord_map(xlim = c(-13, 35),  ylim = c(32, 71))
graph
```

### It can be observed that the Northern part of Europe is more susceptible to crime. The eastern and south western countries have a lower average crime rate. However, Central Europe has a moderate crime average.

## The following graph represents the means of the different types of Crimes
```{r}
value<-c(colMeans(df, na.rm = TRUE))
cols<- colnames(df)
df2<- data.frame(country=cols, value= value)
bplot<- ggplot(df2, aes(x="", y=value, fill=country))+
geom_bar(width = 1, stat = "identity")
bplot
```

### It can be observed that the average number of Theft crimes are the highest and the least occuring crime is Intentional Homicide when compared to the other types of crimes. Moreover, the average rate of Intentional homicide is very minimal and is almost negligible.

